
---
title: "Assignment 9"
author: "Kaitlyn Fales"
date: "10/5/2020"
output: html_document
---


***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(tidyverse)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df1 <- df %>% 
  filter(Date_reported == "2020-10-5") %>% 
  arrange(-Cumulative_deaths)
# The three countries with the highest cumulative deaths are the US, Brazil, and India.
```


2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
library(gganimate)
df %>% 
  filter(Country %in% c("United States of America","Brazil","India")) %>% 
  ggplot(aes(x=Date_reported,y=New_deaths,color=Country)) + geom_line() +
  geom_point(size = 3) + transition_reveal(Date_reported) + 
  labs(x="Date",y="Deaths",title="Deaths over Time")
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df2 <- df %>% 
  filter(Date_reported == "2020-10-5") %>% 
  mutate(death_per_cases = Cumulative_deaths/Cumulative_cases) %>% 
  arrange(-death_per_cases)
# The three countries with the highest death_per_cases as of 10-5-2020 are Yemen, Italy, and Mexico.
```


4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico.
```{r,warning=FALSE,message=FALSE}
df %>% 
  filter(Country %in% c("United States of America","Italy","Mexico")) %>% 
  mutate(death_per_cases = Cumulative_deaths/Cumulative_cases) %>%
  ggplot(aes(x=Date_reported,y=death_per_cases,color=Country)) + geom_line() +
  geom_point(size=3) + transition_reveal(Date_reported) +
  labs(x="Date",y="Deaths per Cases",title="Deaths per Cases over Time")
```



