---
title: "Assignment 4"
author: "Kaitlyn Fales"
date: "9/9/2020"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so. 

```{r, eval=FALSE}
install.packages('tidyverse')
```

-------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://covidtracking.com/data/download/all-states-history.csv).  Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv. 
```{r}
library(tidyverse)
COVID <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```

-------

#### 3.  Fix the date and create some new variables

- Use the following codes to change the type of the `date` column to `date`. 

```{r, eval=FALSE}
library(lubridate)
df$date <- ymd(df$date)
```
```{r}
library(lubridate)
COVID$date <- ymd(COVID$date)
```

- Use the below codes to create `month`, `weekday` and `monthday` variables

```{r, eval=FALSE}

df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```
```{r}
COVID$month <- month(COVID$date)

COVID$weekday <- wday(COVID$date)

COVID$monthday <- mday(COVID$date)
```

-------

#### 4. Create new variables with `case_when`. 

The function `case_when` is a good option to create a new variable from existing variable.  For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable.  `deathIncrease` is the number of daily new death by Covid19.  The new variable `daily_death` takes three values:  low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14).  

```{r, eval=FALSE}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
)
```

- Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month > 20).
```{r}
COVID$month2 <- case_when(
  COVID$monthday <=10 ~ 'early_month',
  COVID$monthday <=20 ~ 'mid_month',
  TRUE ~ 'end_month'
)
```

-------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.
```{r}
COVID <- select(COVID, -totalTestsViral)
```

-------

#### 6. Pipe Operator ( %>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable.  Pipe works very well with all the `tidyverse` packages. Refer to [these slides](file:///C:/Users/sonou/Dropbox/git/math421/slides/dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r, eval=FALSE}
x <- c(1:10)

# square root of x
sqrt(x)

sum(sqrt(x))

log(sum(sqrt(x)))

# log base 2 of 16
log(16, 2)

```
```{r}
x <- c(1:10)

x %>% sqrt
x %>% sqrt %>% sum
x %>% sqrt %>% sum %>% log

16 %>% log(2)
```

-------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

```{r, eval=FALSE}
df %>% 
  group_by(dataQualityGrade) %>% 
  summarise(mean(positiveIncrease))
```

- Calculate the median number of cases (`positiveIncrease`) by `dataQualityGrade`
```{r}
median <- COVID %>% 
  group_by(dataQualityGrade) %>% 
  summarise(median(positiveIncrease))
median
```

- Calculate the maximum number of cases (`positiveIncrease`) by `dataQualityGrade`
```{r}
max <- COVID %>% 
  group_by(dataQualityGrade) %>% 
  summarise(max(positiveIncrease))
max
```

- Calculate the average number of cases (`positiveIncrease`) by `month2`
```{r}
avg_month2 <- COVID %>% 
  group_by(month2) %>% 
  summarise(mean(positiveIncrease))
avg_month2
```

- Calculate the average number of cases (`positiveIncrease`) by `weekday`
```{r}
avg_weekday <- COVID %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
avg_weekday
```

-------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade` in RI, we use:

```{r, eval=FALSE}
df %>% 
  filter(state=='RI')
  group_by(dataQualityGrade) %>% 
  summarise(mean(positiveIncrease))
```

- Calculate the total number of new cases for each weekdays in MA
```{r}
new_cases_weekday <- COVID %>% 
  filter(state=='MA') %>% 
  group_by(weekday) %>% 
  summarise(sum(positiveIncrease))
new_cases_weekday
```

- Calculate the total number of new cases by `month2` in RI and MA.  Is there any difference between the total number of cases for each categories in `month2`? 
```{r}
new_cases_month2 <- COVID %>% 
  filter(state %in% c('MA','RI')) %>% 
  group_by(month2) %>% 
  summarise(sum(positiveIncrease))
new_cases_month2
```

- Compare the average number of new cases between New England and the other states?  
```{r}
avg_New_England <- COVID %>% 
  filter(state %in% c('MA','RI','NH','ME','CT','VE')) %>% 
  summarise(mean(positiveIncrease))
avg_New_England
```
```{r}
avg_other_states <- COVID %>% 
  filter(state != c('MA','RI','NH','ME','CT','VE')) %>% 
  summarise(mean(positiveIncrease))
avg_other_states
```
```{r}
avg_other_states > avg_New_England
```


