
---
title: "Final Project - Building Shiny Apps"
author: "Kaitlyn Fales"
date: "11/24/2020"
output: html_document
---

-------

### Question 1.

Install the following packages `shiny`, `packrat`, `rsconnect`. 

```{r,message=FALSE}
library(shiny)
library(packrat)
library(rsconnect)
```


In this question, you will build and publish your first shiny app. Following these steps to do so

1. Create a new R Script (File -> New R Script). 

2. Copy [these code](https://github.com/bryantstats/math421/blob/master/c1.R) to the file 

3. Save the file as app.R file into a newly created folder. 

4. Press `Ctrl + Shift + Enter` to run the app locally. 

5. Go to https://www.shinyapps.io/

6. Create an account 

7. Go to https://www.shinyapps.io/admin/#/dashboard.  Copy the secret code. 

It should look like this


```{r, eval=FALSE}
rsconnect::setAccountInfo(name='fall20',
			  token='E130D64F22776383660DA7EA5251EC04',
			  secret='mJI7rgq2Wf46g2Wf46g2WNOa+E6NRqr26yG3N5')
```

8. Go back to the app (If you closed it, hit `Ctrl + Shift + Enter` to reopen it), click to `Publish` on the top right corner. 

9.  Paste the secret code and follow the instruction to publish the app on the Internet. 

[Question 1 App](https://kfales.shinyapps.io/Question_1_app/)

-------

### Question 2

Write a shiny app that can plot barplot of two categorical variables in the titanic dataset.

Notice:  Two different shiny apps should be in two different folders. 

[Question 2 App](https://kfales.shinyapps.io/Question_2/)

-------

### Question 3

Write a shiny app that can plot barplot of two categorical variables in an uploaded dataset.

[Question 3 App](https://kfales.shinyapps.io/Question_3/)

-------

### Question 4

Write a shiny app that can plot the scatter plot of two numeric variables and colored by a categorical variable in an uploaded dataset. Example of the plot: https://bryantstats.github.io/math421/slides/6_viz.html#15

Notice: that you can only run 5 apps on shinyapps.io.  If you are running out of apps, you may want to create another account and use multiple accounts.  You may want to try publishing a few times. You could also embed all your shiny apps in an Rmarkdown.  Do the follows to embed an shiny app to an rmarkdown. 

[Question 4 App](https://kfales.shinyapps.io/Question_4/)
  

-------

### Question 5 

Write a shiny app create a plot for the titanic dataset.  The user inputs two variables and can change the range of `Age` (i.e., use `sliderInput`). 

[Question 5 App](https://kfales.shinyapps.io/Question_5/)

-------

### Question 6

Write a shiny app create plots for the titanic dataset. The user inputs two variables and can choose a multiple choice option (i.e., use`radioButtons`). 

[Question 6 App](https://kaitlynfales.shinyapps.io/Question_6/)

-------

### Question 7

Write a shiny app create a plot for the titanic dataset. The user inputs two variables and can choose a tick on some options of a variable (i.e., use`checkboxGroupInput`). 

[Question 7 App](https://kaitlynfales.shinyapps.io/Question_7/)

-------

### Question 8

Write a shiny app create a plot for the titanic dataset. Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput`. 

[Question 8 App](https://kaitlynfales.shinyapps.io/Question_8/)

-------

### Question 9 

Use the below data to write this shiny app: https://fall20.shinyapps.io/app9/.  Check out the app carefully to see its features. 

The data: https://covidtracking.com/data/download/all-states-history.csv

[Question 9 App](https://kaitlynfales.shinyapps.io/Question_9/)

-------

### Question 10 

Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput` to write the second app on the data used on question 9. 

The data: https://covidtracking.com/data/download/all-states-history.csv

[QUestion 10 App](https://kaitlynfales.shinyapps.io/Question_10/)

-------

### Question 11

Write a shiny app on own selected data.  Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput`. 

[Question 11 App](https://kaitlynfales421.shinyapps.io/Question_11/)

-------

### Question 12

Follow this below example to include `navbarPage` to the app in Question 11. You can also write another app that make uses of `selectInput`, `sliderInput`, `radioButtons`, `checkboxGroupInput` and `navbarPage`.

[Question 12 App](https://kaitlynfales421.shinyapps.io/Question_12/)

-------

### Question 13

Present your shiny app in Question 12 on the final exam day. 

-------
