
---
title: "Final Project - Building Shiny Apps"
author: "Kaitlyn Fales"
date: "11/24/2020"
output: html_document
---

-------

### Question 1.

Install the following packages `shiny`, `packrat`, `rsconnect`. 

```{r,message=FALSE,warning=FALSE}
library(shiny)
library(packrat)
library(rsconnect)
```


In this question, you will build and publish your first shiny app. Following these steps to do so

1. Create a new R Script (File -> New R Script). 

2. Copy [these code](https://github.com/bryantstats/math421/blob/master/c1.R) to the file 

3. Save the file as app.R file into a newly created folder. 

4. Press `Ctrl + Shift + Enter` to run the app locally. 

5. Go to https://www.shinyapps.io/

6. Create an account 

7. Go to https://www.shinyapps.io/admin/#/dashboard.  Copy the secret code. 

It should look like this


```{r, eval=FALSE}
rsconnect::setAccountInfo(name='fall20',
			  token='E130D64F22776383660DA7EA5251EC04',
			  secret='mJI7rgq2Wf46g2Wf46g2WNOa+E6NRqr26yG3N5')
```

8. Go back to the app (If you closed it, hit `Ctrl + Shift + Enter` to reopen it), click to `Publish` on the top right corner. 

9.  Paste the secret code and follow the instruction to publish the app on the Internet. 

[Question 1 App](https://kfales.shinyapps.io/Question_1_app/)

```{r,eval=FALSE}
# Set User Interface
ui <- fluidPage(
  titlePanel("Hello Shiny!"),
  
  sidebarLayout(
    
    sidebarPanel(
      sliderInput(
        inputId = "var1",
        label = "Decide a number",
        min = 1,
        max = 100,
        value=10
      )
      
    ),
    
    mainPanel(
      # Output: Histogram ----
      plotOutput(outputId = 'blah1')
    )
    
  )
  
)


######################################
# Main codes for the app
server <- function(input, output) {
  
  output$blah1 <-renderPlot({
    m = input$var1
    hist(rnorm(mean=m, n=1000))
  }
  )
  
  
}

######################################
# Run the app
shinyApp(ui = ui, server = server)
```


-------

### Question 2

Write a shiny app that can plot barplot of two categorical variables in the titanic dataset.

Notice:  Two different shiny apps should be in two different folders. 

[Question 2 App](https://kfales.shinyapps.io/Question_2/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')

variables_names = names(d)

ui <- fluidPage(
  
  titlePanel("Barplots"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Categorical Variable",
        choices = variables_names, selected = "Pclass"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names,
        selected = "Sex"
      )
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    d = read_csv('titanic.csv')
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    r = ggplot(d, aes(x = as.factor(d[[v1]]), fill = as.factor(d[[v2]])))+
      geom_bar()+
      labs(x = v1, fill = v2)
    
    return(r)
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 3

Write a shiny app that can plot barplot of two categorical variables in an uploaded dataset.

[Question 3 App](https://kfales.shinyapps.io/Question_3/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)
library(DT)

ui <- fluidPage(
  
  titlePanel("Barplot Visualization"),
  
  sidebarLayout(
    
    sidebarPanel(
      fileInput('f1', label = 'Upload data for visualization', accept = '.csv'),
      
      selectInput('v1', label='Select a Categorical Variable', choices=''),
      selectInput('v2', label='Select a Categorical Variable', choices='')
      
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server
server <- function(input, output, session) {
  
  myData <- reactive({
    inFile = input$f1
    if (is.null(inFile)) return(NULL)
    data <- read.csv(inFile$datapath, header = TRUE)
    data
  }
  )
  
  output$show_plot <- renderPlot({
    
    inFile = input$f1
    v1 = input$v1
    d <- read.csv(inFile$datapath, header = TRUE)
    
    v1 = input$v1
    v2 = input$v2
    
    
    library(ggplot2)
    
    r = ggplot(d, aes(x = as.factor(d[[v1]]), fill = as.factor(d[[v2]])))+
      geom_bar()+
      labs(x = v1, fill = v2)
    
    return(r)
    
  })
  
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v1', choices = names(data))}
  )
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v2', choices = names(data))}
  )
  
}


shinyApp(ui = ui, server = server)
```


-------

### Question 4

Write a shiny app that can plot the scatter plot of two numeric variables and colored by a categorical variable in an uploaded dataset. Example of the plot: https://bryantstats.github.io/math421/slides/6_viz.html#15

Notice: that you can only run 5 apps on shinyapps.io.  If you are running out of apps, you may want to create another account and use multiple accounts.  You may want to try publishing a few times. You could also embed all your shiny apps in an Rmarkdown.  Do the follows to embed an shiny app to an rmarkdown. 

[Question 4 App](https://kfales.shinyapps.io/Question_4/)
 
```{r,eval=FALSE}
library(tidyverse)
library(shiny)
library(DT)

ui <- fluidPage(
  
  titlePanel("Visualization"),
  
  sidebarLayout(
    
    sidebarPanel(
      fileInput('f1', label = 'Upload data for visualization', accept = '.csv'),
      
      selectInput('v1', label='Select a Numeric Variable', choices=''),
      selectInput('v2', label='Select a Numeric Variable', choices=''),
      selectInput('v3', label='Select a Categorical Variable', choices='')
      
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server
server <- function(input, output, session) {
  
  myData <- reactive({
    inFile = input$f1
    if (is.null(inFile)) return(NULL)
    data <- read.csv(inFile$datapath, header = TRUE)
    data
  }
  )
  
  output$show_plot <- renderPlot({
    
    inFile = input$f1
    v1 = input$v1
    d <- read.csv(inFile$datapath, header = TRUE)
    
    v1 = input$v1
    v2 = input$v2
    v3 = input$v3
    
    
    library(ggplot2)
    
    r = ggplot(d, aes(x = d[[v1]], y = d[[v2]], color = as.factor(d[[v3]])))+
      geom_point()+
      labs(x = v1, y = v2, color = v3)
    
    return(r)
    
  })
  
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v1', choices = names(data))}
  )
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v2', choices = names(data))}
  )
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v3', choices = names(data))}
  )
  
}


shinyApp(ui = ui, server = server)
```
 

-------

### Question 5 

Write a shiny app create a plot for the titanic dataset.  The user inputs two variables and can change the range of `Age` (i.e., use `sliderInput`). 

[Question 5 App](https://kfales.shinyapps.io/Question_5/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')

variables_names = names(d)

ui <- fluidPage(
  
  titlePanel("Bar Plot"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Categorical Variable",
        choices = variables_names, selected = "SibSp"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names,
        selected = "Sex"
      ), 
      
      sliderInput(inputId = "age",
                  "Select Age Range:",
                  min = min(d$Age, na.rm=TRUE),
                  max = max(d$Age, na.rm=TRUE),
                  value= c(1, 80))
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    d <- d %>% filter(Age>input$age[1], Age<input$age[2])
    
    ggplot(d, aes(x = d[[v1]], fill = as.factor(d[[v2]])))+
      geom_bar()+
      labs(x = v1, fill = v2)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 6

Write a shiny app create plots for the titanic dataset. The user inputs two variables and can choose a multiple choice option (i.e., use`radioButtons`). 

[Question 6 App](https://kaitlynfales.shinyapps.io/Question_6/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')

variables_names = names(d)

ui <- fluidPage(
  
  titlePanel("Plots for the Titanic Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_names, selected = "SibSp"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names,
        selected = "Sex"
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Boxplot" = "boxplot",
                               "Density Plot" = "density",
                               "Bar Plot" = "barplot"),
                   selected = 'barplot')
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    if(input$plot_choice == 'boxplot')
    {
      ggplot(d, aes(x = d[[v1]], color = as.factor(d[[v2]])))+
        geom_boxplot()+
        labs(x = v1, color = v2)
    }
    
    else if(input$plot_choice == 'density')
    {
      ggplot(d, aes(x = d[[v1]], color = as.factor(d[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    }
    
    else
    {
      ggplot(d, aes(x = d[[v1]], fill = as.factor(d[[v2]])))+
        geom_bar()+
        labs(x = v1, fill = v2)
    }
    
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 7

Write a shiny app create a plot for the titanic dataset. The user inputs two variables and can choose a tick on some options of a variable (i.e., use`checkboxGroupInput`). 

[Question 7 App](https://kaitlynfales.shinyapps.io/Question_7/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')

variables_names = names(d)


ui <- fluidPage(
  titlePanel("Bar Plot"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        inputId ="var1",
        label = "Select a Categorical Variable",
        choices = variables_names, selected = "Survived"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names,
        selected = "Sex"
      ),
      
      checkboxGroupInput(inputId = "Pclass", label = "Select Passenger Class",
                         choices = names(table(d$Pclass)), inline = TRUE,
                         selected = "1"),
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    d <- d %>% filter(Pclass %in% input$Pclass)
    
    library(ggplot2)
    
    ggplot(d, aes(x = as.factor(d[[v1]]), fill = as.factor(d[[v2]])))+
      geom_bar()+
      labs(x = v1, fill = v2)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 8

Write a shiny app create a plot for the titanic dataset. Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput`. 

[Question 8 App](https://kaitlynfales.shinyapps.io/Question_8/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')

variables_names = names(d)


ui <- fluidPage(
  titlePanel("Plotting the Titanic Data"),
  sidebarLayout(
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_names, selected = "Fare"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names, selected = "Sex"
      ),
      
      sliderInput(inputId = "age",
                  "Select Age Range:",
                  min = min(d$Age, na.rm=TRUE),
                  max = max(d$Age, na.rm=TRUE),
                  value= c(1, 80)
      ),
      
      checkboxGroupInput(inputId = "Pclass", label = "Select Passenger Class",
                         choices = names(table(d$Pclass)), inline = TRUE,
                         selected = "1")
      ,
    
    radioButtons(inputId = "plot_choice", 
                 label = h3("Select Plot:"),
                 choices = c("Boxplot" = "boxplot",
                             "Density Plot" = "density",
                             "Bar Plot" = "barplot"),
                 selected = 'density')
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    d <- d %>% filter(Age>input$age[1], Age<input$age[2])
    d <- d %>% filter(Pclass %in% input$Pclass)
    
    if(input$plot_choice == 'boxplot')
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_boxplot()+
        labs(x = v1, color = v2)
    }
    
    else if(input$plot_choice == 'density')
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    }
    
    else
    {
      ggplot(d, aes(x = (d[[v1]]), fill = as.factor(d[[v2]])))+
        geom_bar()+
        labs(x = v1, fill = v2)
    }
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 9 

Use the below data to write this shiny app: https://fall20.shinyapps.io/app9/.  Check out the app carefully to see its features. 

The data: https://covidtracking.com/data/download/all-states-history.csv

[Question 9 App](https://kaitlynfales.shinyapps.io/Question_9/)

```{r, eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('https://covidtracking.com/data/download/all-states-history.csv')

variables_names = names(d)

d <- d %>% 
  filter(state %in% c('CA','FL','TX','MA','NY','OH'))


ui <- fluidPage(
  titlePanel("Covid 19 by States"),
  sidebarLayout(
    sidebarPanel(
      
      checkboxGroupInput(inputId = "state", label = "Select state to compare",
                         choices = c("CA","FL","TX","MA","NY","OH"), inline = TRUE,
                         selected = "CA")
      ,
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_names, selected = "positive"
      ),
      
      sliderInput(inputId = "Date",
                  "Select Date Range:",
                  min = min(d$date, na.rm = TRUE),
                  max = max(d$date, na.rm = TRUE),
                  value = c(min(d$date),max(d$date)),
                  timeFormat = "%d %b"
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Line Plot" = "line_plot",
                               "Point Plot" = "point_plot"),
                                selected = 'line_plot')
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    
    library(ggplot2)
    
    d <- d %>% filter(date>input$Date[1], date<input$Date[2])
    d <- d %>% filter(state %in% input$state)
    
    if(input$plot_choice == 'line_plot')
    {
      ggplot(d, aes(y = (d[[v1]]), x = (d$date), color = d$state))+
        geom_line()+
        labs(x = "Date", y = v1, color = d$state)
    }
    
    else
    {
      ggplot(d, aes(y = (d[[v1]]), x = (d$date), color = d$state))+
        geom_point()+
        labs(x = "Date", y = v1, color = d$state)
    }
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 10 

Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput` to write the second app on the data used on question 9. 

The data: https://covidtracking.com/data/download/all-states-history.csv

[Question 10 App](https://kaitlynfales.shinyapps.io/Question_10/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('https://covidtracking.com/data/download/all-states-history.csv')

variables_names = names(d)


ui <- fluidPage(
  titlePanel("Covid 19"),
  sidebarLayout(
    sidebarPanel(
      
      checkboxGroupInput(inputId = "dataquality", label = "Select Data Quality Grade",
                         choices = c("A+","A","B","C","D","F"), inline = TRUE,
                         selected = c("A+","A"))
      ,
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_names, selected = "recovered"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a State",
        choices = names(table(d$state)), selected = "MA"
      ),
      
      sliderInput(inputId = "Date",
                  "Select Date Range:",
                  min = min(d$date, na.rm = TRUE),
                  max = max(d$date, na.rm = TRUE),
                  value = c(min(d$date),max(d$date)),
                  timeFormat = "%d %b"
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Line Plot" = "line_plot",
                               "Point Plot" = "point_plot",
                               "Bar Plot" = "bar_plot"),
                   selected = 'line_plot')
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    
    library(ggplot2)
    
    d <- d %>% filter(date>input$Date[1], date<input$Date[2])
    d <- d %>% filter(dataQualityGrade %in% input$dataquality)
    d <- d %>% filter(state %in% input$var2)
    
    if(input$plot_choice == 'line_plot')
    {
      ggplot(d, aes(y = (d[[v1]]), x = (d$date)))+
        geom_line()+
        labs(x = "Date", y = v1)
    }
    
    else if(input$plot_choice == 'bar_plot')
    {
      ggplot(d, aes(y = (d[[v1]]), x = (d$dataQualityGrade)))+
        geom_col()+
        labs(x = "Data Quality Grade", y = v1)
    }
    
    else
    {
      ggplot(d, aes(y = (d[[v1]]), x = (d$date)))+
        geom_point()+
        labs(x = "Date", y = v1)
    }
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 11

Write a shiny app on own selected data.  Make uses of `selectInput`, `sliderInput`, `radioButtons`, and `checkboxGroupInput`. 

[Question 11 App](https://kaitlynfales421.shinyapps.io/Question_11/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('insurance.csv')

variables_names = names(d)


ui <- fluidPage(
  titlePanel("Plotting Health Insurance Data"),
  sidebarLayout(
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_names, selected = "charges"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_names, selected = "sex"
      ),
      
      sliderInput(inputId = "bmi",
                  "Select BMI Range:",
                  min = min(d$bmi, na.rm=TRUE),
                  max = max(d$bmi, na.rm=TRUE),
                  value= c(min(d$bmi),max(d$bmi))
      ),
      
      checkboxGroupInput(inputId = "Region", label = "Select Region",
                         choices = names(table(d$region)), inline = TRUE,
                         selected = c("northeast","northwest")
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Boxplot" = "boxplot",
                               "Density Plot" = "density"),
                   selected = 'boxplot')
    ),
    mainPanel(
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    d <- d %>% filter(bmi>input$bmi[1], bmi<input$bmi[2])
    d <- d %>% filter(region %in% input$Region)
    
    if(input$plot_choice == 'boxplot')
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_boxplot()+
        labs(x = v1, color = v2)
    }
    
    else
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    }
    
    
  })

}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 12

Follow this below example to include `navbarPage` to the app in Question 11. You can also write another app that make uses of `selectInput`, `sliderInput`, `radioButtons`, `checkboxGroupInput` and `navbarPage`.

[Question 12 App](https://kaitlynfales421.shinyapps.io/Question_12/)

```{r,eval=FALSE}
library(tidyverse)
library(shiny)

d = read_csv('insurance.csv')

variables_names = names(d)


ui <- navbarPage("Navigation",
                 tabPanel("Numeric and Categorical Plots",
                          titlePanel("Plotting Health Insurance Data"),
                          sidebarLayout(
                            sidebarPanel(
                              
                              selectInput(
                                inputId ="var1",
                                label = "Select a Numeric Variable",
                                choices = variables_names, selected = "charges"
                              ),
                              
                              selectInput(
                                inputId ="var2",
                                label = "Select a Categorical Variable",
                                choices = variables_names, selected = "sex"
                              ),
                              
                              sliderInput(inputId = "bmi",
                                          "Select BMI Range:",
                                          min = min(d$bmi, na.rm=TRUE),
                                          max = max(d$bmi, na.rm=TRUE),
                                          value= c(min(d$bmi),max(d$bmi))
                              ),
                              
                              checkboxGroupInput(inputId = "Region", label = "Select Region",
                                                 choices = names(table(d$region)), inline = TRUE,
                                                 selected = c("northeast","northwest")
                              ),
                              
                              radioButtons(inputId = "plot_choice", 
                                           label = h3("Select Plot:"),
                                           choices = c("Boxplot" = "boxplot",
                                                       "Density Plot" = "density"),
                                           selected = 'boxplot')
                            ),
                            mainPanel(
                              
                              plotOutput(outputId = 'show_plot')
                            )
                          )
                 ),
                 
                 tabPanel("Bar Plots",
                          titlePanel("Plotting Health Insurance Data"),
                          sidebarLayout(
                            sidebarPanel(
                              
                              selectInput(
                                inputId ="var3",
                                label = "Select a Categorical Variable",
                                choices = variables_names, selected = "children"
                              ),
                              
                              selectInput(
                                inputId ="var4",
                                label = "Select a Categorical Variable",
                                choices = variables_names, selected = "smoker"
                              ),
                              
                              sliderInput(inputId = "bmi1",
                                          "Select BMI Range:",
                                          min = min(d$bmi, na.rm=TRUE),
                                          max = max(d$bmi, na.rm=TRUE),
                                          value= c(min(d$bmi),max(d$bmi))
                              ),
                              
                              checkboxGroupInput(inputId = "Region1", label = "Select Region",
                                                 choices = names(table(d$region)), inline = TRUE,
                                                 selected = c("northeast","northwest"))
                            ),
                            
                            mainPanel(
                              
                              plotOutput(outputId = 'show_plot2')
                            )
                          )
                 )
)


# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    
    
    library(ggplot2)
    
    d <- d %>% filter(bmi>input$bmi[1], bmi<input$bmi[2])
    d <- d %>% filter(region %in% input$Region)
    
    if(input$plot_choice == 'boxplot')
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_boxplot()+
        labs(x = v1, color = v2)
    }
    
    else
    {
      ggplot(d, aes(x = (d[[v1]]), color = as.factor(d[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    }
    
    
  })
  
  output$show_plot2 <- renderPlot({
    
    v3 = input$var3
    v4 = input$var4
    
    
    library(ggplot2)
    
    d <- d %>% filter(bmi>input$bmi1[1], bmi<input$bmi1[2])
    d <- d %>% filter(region %in% input$Region1)
    
    ggplot(d, aes(x = as.factor(d[[v3]]), fill = as.factor(d[[v4]]))) + 
      geom_bar() + labs(x = v3, fill = v4)
    
    
  })
  
  
}
# app
shinyApp(ui = ui, server = server)
```


-------

### Question 13

Present your shiny app in Question 12 on the final exam day. 

-------
